---
title: "Statistical Inference - Course Project"
author: "Ken Wood"
date: "8/17/2020"
output: html_document
---

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with the expression `rexp(n, lambda)` where `lambda` is the rate parameter. The mean of the exponential distribution is `1/lambda` and the standard deviation is also `1/lambda`. We will set $\lambda$ = 0.2 for all of the simulations. We will investigate the` distribution of averages of 40 exponentials. Note that we will need to do a thousand simulations.

Our goal is to illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. We will: 

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

In deliverable 3, we will focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

First, let's plot the exponential distribution for 1000 observations using $\lambda$ = 0.2. The theoretical mean and standard deviation for this distribution are $\mu$ and $\sigma$, respectively and are both equal to 1/$\lambda$ = 5.

If we simulate 1000 observations of the exponential distribution using the `rexp` function in R, we get the following histogram with sample mean and standard deviation values very close to 5.

```{r}
exp_dist <- rexp(1000,rate = 0.2)
hist(exp_dist,breaks=20,col="blue",xlab = "x",main="Histogram of an Exponential Distribution - 1000 Observations")
mean(exp_dist)
sd(exp_dist)
```

```{r}
mnsexp = NULL
for (i in 1 : 1000) mnsexp = c(mnsexp, mean(rexp(40,rate=0.2)))
hist(mnsexp)
mean(mnsexp)
sd(mnsexp)
```
Recall that the mean and standard deviation for an exponential distribution = `1/lambda`. If we select `lambda = 0.2`, then the mean $\mu$ and standard deviation $\sigma$ is 5.
