---
title: "Week 4 Programming Assignment"
author: "Ken Wood"
date: "8/2/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}

# Load the raw data files.
# These lines of code will take a little time to execute, so please be patient!

NEI <- readRDS("exdata-data-NEI_data/summarySCC_PM25.rds")
SCC <- readRDS("exdata-data-NEI_data/Source_Classification_Code.rds")
merged_df <- merge(NEI,SCC,by="SCC")
```

# Questions

We will address the following questions and tasks in our exploratory analysis. For each question/task we will need to make a single plot. Unless specified, we can use any plotting system in R to make our plot.

1. Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the **base** plotting system, make a plot showing the *total* PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r}

total_emissions <- aggregate(NEI$Emission, by=list(NEI$year), sum)
plot(total_emissions,pch=16,xlab="Year",ylab="Emissions (tons)",main="Total Emissions by Year")
lines(total_emissions$Group.1,total_emissions$x)
grid(lty="dotted")
```

2. Have total emissions from PM2.5 decreased in the **Baltimore City, Maryland** (`fips == "24510"`) from 1999 to 2008? Use the base plotting system to make a plot answering this question.

```{r}

# subset NEI for Baltimore City (fips=="24510")
baltimore_city <- subset(NEI,fips=="24510")
balt_emissions <- aggregate(baltimore_city$Emission, by=list(baltimore_city$year), sum)

plot(balt_emissions,pch=16,xlab="Year",ylab="Emissions (tons)",main="Total Emissions by Year: Baltimore-City")
lines(balt_emissions$Group.1,balt_emissions$x)
grid(lty="dotted")
```

3. Of the four types of sources indicated by the `type` (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for **Baltimore City**? Which have seen increases in emissions from 1999–2008? Use the **ggplot2** plotting system to make a plot answer this question.

```{r}

library(ggplot2)

balt_emissions_type_year <- aggregate(baltimore_city$Emission, by=list(baltimore_city$type,baltimore_city$year), sum)
names(balt_emissions_type_year) <- c("type","Year","Emissions")

ggplot(balt_emissions_type_year, aes(Year,Emissions))+ geom_point(aes(color=type))+geom_line(aes(color=type))+ggtitle("Total Emissions by Type and Year: Baltimore-City")+ylab("Emissions (tons)")+labs(color="Emission Type")
```

4. Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

```{r}
library(dplyr)
library(ggplot2)

# Filter the merged df for 'coal combustion-related' sources
coal_combustion <- dplyr::filter(merged_df, grepl('Coal', Short.Name) & grepl('Ext Comb',Short.Name))
emissions_type_year <- aggregate(coal_combustion$Emission, by=list(coal_combustion$type,coal_combustion$year), sum)
names(emissions_type_year) <- c("type","Year","Emissions")

ggplot(emissions_type_year, aes(Year,Emissions))+ geom_point(aes(color=type))+geom_line(aes(color=type))+ggtitle("Total Coal Combustible-Related Emissions by Type and Year")+ylab("Emissions (tons)")+labs(color="Emission Type")
```

5. How have emissions from motor vehicle sources changed from 1999–2008 in **Baltimore City**?

```{r}

library(dplyr)
library(ggplot2)

# Filter the merged_df for 'Baltimore City' and 'Highway Veh'
balt_cars <- dplyr::filter(merged_df, grepl("24510",fips) & grepl('Highway Veh',Short.Name))
balt_car_emissions <- aggregate(balt_cars$Emission, by=list(balt_cars$type,balt_cars$year), sum)
names(balt_car_emissions) <- c("type","Year","Emissions")

ggplot(balt_car_emissions, aes(Year,Emissions))+ geom_point(aes(color=type))+geom_line(aes(color=type))+ggtitle("Total Motor Vehicle Emissions by Type and Year: Baltimore-City")+ylab("Emissions (tons)")+labs(color="Emission Type")
```

6. Compare emissions from motor vehicle sources in **Baltimore City** with emissions from motor vehicle sources in **Los Angeles County, California** (`fips == "06037"`). Which city has seen greater changes over time in motor vehicle emissions?

```{r}
library(dplyr)
library(ggplot2)

# Filter the merged_df for fips==24510|fips==06037 and 'Highway Veh'
balt_la_cars <- dplyr::filter(merged_df, grepl("24510|06037",fips) & grepl('Highway Veh',Short.Name))
balt_la_car_emissions <- aggregate(balt_la_cars$Emission, by=list(balt_la_cars$fips,balt_la_cars$year), sum)
names(balt_la_car_emissions) <- c("fips","Year","Emissions")

ggplot(balt_la_car_emissions, aes(Year,Emissions))+ geom_point(aes(color=fips))+geom_line(aes(color=fips))+ggtitle("Total Motor Vehicle Emissions by FIPS Location and Year")+ylab("Emissions (tons)")+labs(color="fips location")
```



